---
title: "Problem Set 2 - Demonstrate the convergence of a t distribution"
output: html_notebook
---

# Group 3 : Sai Myint, Yoojin Oh, Liz Ward, Karl Wirth

```{r}
library(ggplot2)
library(cowplot)
```

## Question 1:

### Generate 50,000 observations from a random t (gen t03= rt(3)) distribution with the following degrees of freedom [3, 6, 12, 27, 59] and also a standard normal distribution.

\
Setting up the plots for the t distributions and standard normal

```{r}
# Creating t distributions for each degree of freedom
t_dist3 <- rt(50000,3)
t_dist6 <- rt(50000,6)
t_dist12 <- rt(50000,12)
t_dist27 <- rt(50000,27)
t_dist59 <- rt(50000,59)

t_distdf <- data.frame(t_dist3,t_dist6,t_dist12,t_dist27,t_dist59)
# Distribution for standard normal
normal <- rnorm(50000)
```

Generating the Plots for t distribution [3,6,12,27,59]

```{r}
hist(t_dist3)
hist(t_dist6)
hist(t_dist12)
hist(t_dist27)
hist(t_dist59)
hist(normal)
```

We can see that when the degrees of freedom increases in the t-distribution it approaches a standard normal. With the degree of freedom at 59, the distribution looks very close to a standard normal distribution.

## Two Tailed 95% Scores of the T Distribution

Determine the two tailed 95% scores of the t distribution for each degree of freedom and compare them to a z of the significance (1.96)

#### Calculating the Two Tail 95% score

```{r}
# Z Score for the significance of 1.96
zscore <- 1.96

# Calculating T scores for each Degrees of Freedom
t_scores <- qt(0.95, c(3, 6, 12, 27, 59))

# Printing Scores
print (t_scores)
print (zscore)
```

The Z score is significantly higher than the t scores of every distribution. We expected this because the degrees of freedom is like a safety net, and is a more conservative test than the standard normal distribution.

## Kernel Density Plot

```{r}
# Creating Kernel Density Plot for each t-distribution
ggplot(t_distdf,aes(t_dist3)) + geom_histogram(aes(y = stat(density))) + geom_density(col = "red") + ggtitle("T Distribution DF = 3")

ggplot(t_distdf,aes(t_dist6)) + geom_histogram(aes(y = stat(density))) + geom_density(col = "blue") + ggtitle("T Distribution DF = 6")

ggplot(t_distdf,aes(t_dist12)) + geom_histogram(aes(y = stat(density))) + geom_density(col = "green") + ggtitle("T Distribution DF = 12")

ggplot(t_distdf,aes(t_dist27)) + geom_histogram(aes(y = stat(density))) + geom_density(col = "orange") + ggtitle("T Distribution DF = 27")

ggplot(t_distdf,aes(t_dist59)) + geom_histogram(aes(y = stat(density))) + geom_density(col = "yellow") + ggtitle("T Distribution DF = 59")
```

#### (a) At what value do you feel the normal approximation of a t is \"good enough\" and why based on this plot?

Based on the plots, the t-distribution with a degree of freedom of 59 closely resembles a normal distribution. Good enough without a near perfect would be DF=27. The difference between 59 and 27 is that at DF=29 you get more data in the tails than DF=27.

#### (a) Why would a normal distribution be a poor approximation for a t-distribution with 3 degrees of freedom. Make a normal probability plot (pnorm t03) and interpret it as part of your answer.

The distribution for t-distribution with 3 DF is very peaked compared to the normal probability plot. This means that there are more extreme values in the t-distribution compared to the standard normal. In the standard normal plot, there is even distribution within the tails and throughout the curve.

We can see from the graph below, that there are going to be missing values (ie. the gaps).

```{r}
norm <- rnorm(50000)
normDF <- data.frame(norm,t_dist3)

plot(density(normDF$norm))
lines(density(normDF$t_dist3))
```
